---
title: "Strings and Factors"
output: github_document
date: "2024-10-28"
---

```{r setup}

library(tidyverse)

library(p8105.datasets)
library (rvest)

knitr::opts_chunk$set(
  fig.width = 6, 
  fig.asp = 0.6, 
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))
          
options(
  ggplot2.continuous.colour = "viridis", 
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Strings and regex

Okay! So a string is a vector of characters.That's the best definition! You can use the str_detect function to find a pattern between the character and other components of the vector. Also make sure you run the libraries above so the functions below work! For example, below, the result told me hey, there's a jeff in the 4th position. Regex just stands for regular expression. 

```{r}
string_vec = c("my", "name", "is", "jeff")
str_detect(string_vec, "jeff")
```

I can also search for the letter "e" in the vector. There are 2 characters in the vector where this is the case and that is what it is putting out. 

```{r}
string_vec = c("my", "name", "is", "jeff")
str_detect(string_vec, "e")
```

Another try. This time it's all false because there is no capital J! 
```{r}
string_vec = c("my", "name", "is", "jeff")
str_detect(string_vec, "J")
```


You can also replace components of the string. You enter the first character you have and then the character that you want to replace it with! If I wanted to remove that character, I could just put and empty quotes "" as the argument in the replacement column

```{r}
string_vec = c("my", "name", "is", "jeff")
str_replace(string_vec, "jeff", "Jeff")
```

You can also detect phrases in different locations. 

```{r}
string_vec = c(
  "i think we all rule for participating",
  "i think i have been caught",
  "i think this will be quite fun actually",
  "it will be fun, i think"
  )
str_detect(string_vec, "i think")
```

What if I only want phrases that START with "i think"? I can specify that as well. The carrot says only pick phrases that start with this value. 

```{r}
string_vec = c(
  "i think we all rule for participating",
  "i think i have been caught",
  "i think this will be quite fun actually",
  "it will be fun, i think"
  )
str_detect(string_vec, "^i think")
```

I can also check for the case where "i think" is at the end of the string. 

```{r}
string_vec = c(
  "i think we all rule for participating",
  "i think i have been caught",
  "i think this will be quite fun actually",
  "it will be fun, i think"
  )
str_detect(string_vec, "i think$")
```

We have another example. Here, we see that there are many expressions with the word Bush in it. BUT since the `str_detect` function is case sensitive, how do I detect all versions? Use brackets saying what the options are for the first letter. 

```{r}
string_vec = c("Y'all remember Pres. HW Bush?", 
               "I saw a green bush", 
               "BBQ and Bushwalking at Molongo Gorge", 
               "BUSH -- LIVE IN CONCERT!!")
str_detect(string_vec, "[Bb]ush")
```

I can also code for more complicated cases. Here, I tell the code to find all cases where I start with a number and then have a letter adjacent. I list the range of numbers and then the range of letters and specify whether it's capitalized or not.

```{r}
string_vec = c(
  '7th inning stretch',
  '1st half soon to begin. Texas won the toss.',
  'she is 5 feet 4 inches tall',
  '3AM - cant sleep :('
  )

str_detect(string_vec, "[0-9][a-zA-Z]")
```

This case is a bit confusing because here the syntax for r conflicts with some of the characters in the code (it could look misleading), but the code below is written to detect cases where there is 7, an adjacent character, and then a 1. That's how r is reading what I've typed. And that's why it returns true true false true. 

Alternatively, I could try to enter the brackets used above. 

```{r}
string_vec = c(
  'Its 7:11 in the evening',
  'want to go to 7-11?',
  'my flight is AA711',
  'NetBios: scanning ip 203.167.114.66'
  )

str_detect(string_vec, "7.11")
```

But what if I wanted 7.11? \\ says treat this dot like a dot, not a place holder for a special character like above. 

```{r}
string_vec = c(
  'Its 7:11 in the evening',
  'want to go to 7-11?',
  'my flight is AA711',
  'NetBios: scanning ip 203.167.114.66'
  )

str_detect(string_vec, "7\\.11")
```

## Factors

So the `factor()` function takes a vector of characters and assigns levels(aka numbers) to it based on the alphabetical order of those characters. It placed female first because it starts with f. But it also assigns numbers based on this order using the `as.numeric` function. 

```{r}
vec_sex = factor(c("male", "male", "female", "female"))
vec_sex
as.numeric(vec_sex)
```

I can also tell R to make the first level to be "male". 

```{r}
vec_sex = factor(c("male", "male", "female", "female"))
vec_sex = fct_relevel(vec_sex, "male")
vec_sex
as.numeric(vec_sex)
```

## NSDUH

This is an online database from the national survey on drug use and health comparison. 
Wow I was finally able to extract the first table from this data set! If you look at the set, the 12+, 12-17, 18-25 corresponds to the age range. 

```{r}
url = "http://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.htm"

table_marj = 
  read_html(url) |> 
  html_table() |> 
  first() |>
  slice(-1)

table_marj
```

You can also remove columns. Here I am removing the P value column.

```{r}
data_marj = 
  table_marj |>
  select(-contains("P Value")) 
data_marj
```
 
 Remember the pivot function? We can use that to reorganize the columns so it's more palatable and I'm going to organize everything that's not state (hence the -state), and I want the column name to be age_year because that's how it's titled. 
 
```{r}
data_marj = 
  table_marj |>
  select(-contains("P Value")) |>
  pivot_longer(
    -State,
    names_to = "age_year", 
    values_to = "percent")
data_marj
```
 
 Now let's say I want to separate the age_year column a little. It actually made that one column into 2 columns, and I told it to separate when it got to the open parenthesis. 
 
```{r}
data_marj = 
  table_marj |>
  select(-contains("P Value")) |>
  pivot_longer(
    -State,
    names_to = "age_year", 
    values_to = "percent") |>
  separate(age_year, into = c("age", "year"), sep = "\\(")
data_marj
```
 
 I can also code to get rid of the closed parenthesis by telling R to replace it with nothing. 

```{r}
data_marj = 
  table_marj |>
  select(-contains("P Value")) |>
  pivot_longer(
    -State,
    names_to = "age_year", 
    values_to = "percent") |>
  separate(age_year, into = c("age", "year"), sep = "\\(") |>
 mutate(
    year = str_replace(year, "\\)", ""))

data_marj
```

Let's do the same for the percent column. I want to get rid of those extra letters. Recall, putting the dollar sign means anything that ends with that character. (and carrot means starts with). I can also simultaneously change the percent column to be a number (double) instead of a character

```{r}
data_marj = 
  table_marj |>
  select(-contains("P Value")) |>
  pivot_longer(
    -State,
    names_to = "age_year", 
    values_to = "percent") |>
  separate(age_year, into = c("age", "year"), sep = "\\(") |>
  mutate(
    year = str_replace(year, "\\)", ""),
    percent = str_replace(percent, "[a-c]$", ""), 
    percent = as.numeric(c(percent))
  )
data_marj
```

Now! Let's say I want to get rid of the Total U.S. rows and only have the state-specific data. Use the filter function- you enter the `!` syntax to say take this out, and we are accessing the state column, and within those columns, you want to filter out the following row of characters.

```{r}
data_marj = 
  table_marj |>
  select(-contains("P Value")) |>
  pivot_longer(
    -State,
    names_to = "age_year", 
    values_to = "percent") |>
  separate(age_year, into = c("age", "year"), sep = "\\(") |>
  mutate(
    year = str_replace(year, "\\)", ""),
    percent = str_replace(percent, "[a-c]$", ""),
    percent = as.numeric(percent)) |>
  filter(!(State %in% c("Total U.S.", "Northeast", "Midwest", "South", "West")))

data_marj
```

You can also use the filter function to INCLUDE only certain information.

```{r}
data_marj =
  data_marj %>% 
  filter(age == "12-17") 
data_marj
```
Now, the data set lists (and would plot) the state variable in alphabetical order. If I wanted to change the display order so it reflected the percentage, as opposed to the name, I could use the factor reorder function, telling it to reorganize state by percent in the plot. 

```{r}
data_marj =
  data_marj %>% 
  filter(age == "12-17") %>% 
  mutate(State = fct_reorder(State, percent)) %>% 
ggplot(aes(x = State, y = percent, color = year)) + 
    geom_point() + 
    theme(axis.text.x = element_text(angle = 90, hjust = 1))

data_marj
```

