---
title: "Strings and Factors"
output: github_document
date: "2024-10-28"
---

```{r setup, include = FALSE}

library(tidyverse)

library(p8105.datasets)
library (rvest)

knitr::opts_chunk$set(
  fig.width = 6, 
  fig.asp = 0.6, 
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))
          
options(
  ggplot2.continuous.colour = "viridis", 
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Strings and regex

Okay! So a string is a vector of characters.That's the best definition! You can use the str_detect function to find a pattern between the character and other components of the vector. Also make sure you run the libraries above so the functions below work! For example, below, the result told me hey, there's a jeff in the 4th position. Regex just stands for regular expression. 

```{r}
string_vec = c("my", "name", "is", "jeff")
str_detect(string_vec, "jeff")
```

I can also search for the letter "e" in the vector. There are 2 characters in the vector where this is the case and that is what it is putting out. 

```{r}
string_vec = c("my", "name", "is", "jeff")
str_detect(string_vec, "e")
```

Another try. This time it's all false because there is no capital J! 
```{r}
string_vec = c("my", "name", "is", "jeff")
str_detect(string_vec, "J")
```


You can also replace components of the string. You enter the first character you have and then the character that you want to replace it with! If I wanted to remove that character, I could just put and empty quotes "" as the argument in the replacement column

```{r}
string_vec = c("my", "name", "is", "jeff")
str_replace(string_vec, "jeff", "Jeff")
```

You can also detect phrases in different locations. 

```{r}
string_vec = c(
  "i think we all rule for participating",
  "i think i have been caught",
  "i think this will be quite fun actually",
  "it will be fun, i think"
  )
str_detect(string_vec, "i think")
```

What if I only want phrases that START with "i think"? I can specify that as well. The carrot says only pick phrases that start with this value. 

```{r}
string_vec = c(
  "i think we all rule for participating",
  "i think i have been caught",
  "i think this will be quite fun actually",
  "it will be fun, i think"
  )
str_detect(string_vec, "^i think")
```

I can also check for the case where "i think" is at the end of the string. 

```{r}
string_vec = c(
  "i think we all rule for participating",
  "i think i have been caught",
  "i think this will be quite fun actually",
  "it will be fun, i think"
  )
str_detect(string_vec, "i think$")
```

We have another example. Here, we see that there are many expressions with the word Bush in it. BUT since the `str_detect` function is case sensitive, how do I detect all versions? Use brackets saying what the options are for the first letter. 

```{r}
string_vec = c("Y'all remember Pres. HW Bush?", 
               "I saw a green bush", 
               "BBQ and Bushwalking at Molongo Gorge", 
               "BUSH -- LIVE IN CONCERT!!")
str_detect(string_vec, "[Bb]ush")
```

I can also code for more complicated cases. Here, I tell the code to find all cases where I start with a number and then have a letter adjacent. I list the range of numbers and then the range of letters and specify whether it's capitalized or not.

```{r}
string_vec = c(
  '7th inning stretch',
  '1st half soon to begin. Texas won the toss.',
  'she is 5 feet 4 inches tall',
  '3AM - cant sleep :('
  )

str_detect(string_vec, "[0-9][a-zA-Z]")
```

This case is a bit confusing because here the syntax for r conflicts with some of the characters in the code (it could look misleading), but the code below is written to detect cases where there is 7, an adjacent character, and then a 1. That's how r is reading what I've typed. And that's why it returns true true false true. 

Alternatively, I could try to enter the brackets used above. 

```{r}
string_vec = c(
  'Its 7:11 in the evening',
  'want to go to 7-11?',
  'my flight is AA711',
  'NetBios: scanning ip 203.167.114.66'
  )

str_detect(string_vec, "7.11")
```

But what if I wanted 7.11? \\ says treat this dot like a dot, not a place holder for a special character like above. 

```{r}
string_vec = c(
  'Its 7:11 in the evening',
  'want to go to 7-11?',
  'my flight is AA711',
  'NetBios: scanning ip 203.167.114.66'
  )

str_detect(string_vec, "7\\.11")
```

## Factors

So the `factor()` function takes a vector of characters and assigns levels(aka numbers) to it based on the alphabetical order of those characters. It placed female first because it starts with f. But it also assigns numbers based on this order using the `as.numeric` function. 

```{r}
vec_sex = factor(c("male", "male", "female", "female"))
vec_sex
as.numeric(vec_sex)
```

I can also tell R to make the first level to be "male". 

```{r}
vec_sex = factor(c("male", "male", "female", "female"))
vec_sex = fct_relevel(vec_sex, "male")
vec_sex
as.numeric(vec_sex)
```

## NSDUH
